---
title: "PA1_template.Rmd"
author: "Christian Llano"
date: "Friday, June 12, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


##Loading and preprocessing the data

I read data using Rstudio option, import data sets.
```{r,echo=TRUE}
activity <- read.csv("activity.csv")
#changing to date type
activity$date<-as.Date(activity$date,"%Y-%m-%d")
head(activity)
```

## What is mean total number of steps taken per day?


```{r,echo=TRUE}
#1. Calculate the total number of steps taken per day
aggregate(steps~date,data=activity,sum,na.rm=T)->x
head(x)

#2 Here a boxplot of x 
barplot(x[,2])

#3 Finding mean total number of steps.

mean(activity$steps,na.rm=TRUE)
#3 Calculating mean and median of the total number of steps taken per day
aggregate(steps~date,data=activity,function(x) c(Mean=mean(x,na.rm=T),Median=median(x,na.rm=T)))


#4 Mean of steps by interval
as.data.frame(aggregate(steps~interval,data=activity,mean,na.rm=TRUE))->x1

#5 Next we obtain the maximum of mean steps by interval
x1[which.max(x1$steps),]
```

We seee that median is 0 in all cases, this happens because most of the steps during intervals resulting to be 0

## What is the average daily activity pattern?

Using the x1 created before, where we summarize mean of steps by intervals

```{r,echo=TRUE}
plot(x1,type="l")
```


## Imputing missing values


```{r,echo=TRUE}

#How many NA values are by column on activity
apply(activity,2,function(x)sum(is.na(x)))

#Looking for mean according to date
x2<-aggregate(steps~date,data=activity,mean,na.rm=TRUE)

#Looking for mean according to intervals
x3<-aggregate(steps~interval,data=activity,mean,na.rm=TRUE)
```

I decided to replace NA values by the mean of its interval as showing next:

```{r,echo=TRUE}
#merging activity and x3 by the column interval.
merge(activity,x3,by = "interval")->data1
head(data1)

#copying values of means by interval to its corresponding NA values
data1$steps.x[is.na(data1$steps.x)]<-data1$steps.y[is.na(data1$steps.x)]

#ordering and looking to first cases
data1<-data1[order(data1$date,data1$interval),]
data1$steps.y<-NULL
names(data1)[2]<-"steps"
head(data1)


#total steps, mean and median by day on data1

table2<-aggregate(steps~date,data=data1,sum)
barplot(table2[,2])
hist(table2$steps)

table3<-aggregate(steps~date,data=data1,function(x) c(Sum= sum(x),Mean=mean(x),Median=median(x)))
head(table3)

```


## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE}

#Creating week according to weekday o weekend on date
data1$week<-ifelse(weekdays(data1$date)%in%c("sábado","domingo"),"weekend","weekday")
head(data1)
table4<-aggregate(steps~interval+week,data=data1,mean)
head(table4)
levels(as.factor(as.character(data1$interval2)))->xlab
```

Now I plot using ggplot2
```{r,echo=TRUE}
### Graphics

library(ggplot2)

ggplot(table4,aes(interval,steps))->qpl1

summary(qpl1)

qpl1+
    geom_line()+
    facet_grid(week~.)+
    xlab("Interval")+
    ylab("Number of steps")
```

We can note there is a remarkable difference onpatterns of steps, in intervals between 10:00 to 20:00 

